---
title: "Lab 10 - Nested Designs"
author: "Drew Shives"
date: "2022-01-26"
slug: lab-10-nested-designs
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2022-04-27T13:31:58-05:00'
featured: no
type: book
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
library(tidyverse)
```

## (1) From Chapter 20, reproduce the bat and hat example (20.2) in R. Your code should represent the data in long-form, conduct the ANOVA, and report the ANOVA table. You will know if you did it correctly if you can reproduce the ANOVA table from the textbook. 

```{r}
subject <- c(rep(c("s1","s2","s3","s4","s5"),2),
             rep(c("s6","s7","s8","s9","s10"),2))
phon_sim <- rep(rep(c("b1","b2"), each=5),2)
age <- rep(c("a1","a2"), each=10)
correct <- c(15,23,12,16,14,
                 13,19,10,16,12,
                 39,31,40,32,38,
                 29,15,30,26,30)

df_bh <- tibble(subject,
                       phon_sim,
                       age,
                       correct)

# run ANOVA
aov_output <- aov(correct ~ phon_sim*age + Error(subject/phon_sim), data = df_bh)

summary(aov_output)
```

## (2) From Chapter 21, reproduce the phonological similarity example (21.2.1) in R. Your code should represent the data in long-form, conduct the ANOVA, and report the ANOVA table. You will know if you did it correctly if you can reproduce the ANOVA table from the textbook.

```{r}
subject <- rep(c("s1","s2","s3","s4"), each=10)
typ_b <- rep(rep(c("b1","b2"), each=5),4)
faces_a <- rep(c("a1","a2","a3","a4","a5"),8)
centiseconds <- c(20,22,25,24,19,37,37,43,48,45,
                   9,8,21,21,21,34,35,35,37,39,
                   18,20,18,21,33,35,39,39,37,40,
                   5,14,16,22,23,38,49,51,50,52)

df_face <-  tibble(subject,
                     typ_b,
                     faces_a,
                     centiseconds)

aov_output <- aov(centiseconds  ~ typ_b + (faces_a%in%typ_b) +
                       Error(subject + subject:faces_a + subject:typ_b),
               data = df_face)

summary(aov_output)

```














