---
title: "Lab 5 - ANOVA and Randomization"
author: Drew Shives
date: '2022-01-31'
slug: lab-5-anova-randomization
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2022-03-10T13:31:58-05:00'
featured: no
type: book
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(kableExtra)

set.seed(1)
```

## Consider a design with 3 groups, and 10 people per group. Assume that the dependent variable is assumed to be normally distributed, and use unit normal distributions with mean = 0, and sd = 1 in your simulations.

### Setting the parameters for the data and subsequent simulations.

```{r}
levels <- 3
n_per_level <- 10
n_samples <- 1000
```

### Creating the null distribution.

```{r}
gen_sim_f <- function(num_levels, num_per_level, num_samples) {
  f_values <- length(num_samples)

  for(i in 1:num_samples) {
    sim_data <- tibble(subject = 1:(num_levels * num_per_level),
                       level = as.factor(rep(1:num_levels, each = num_per_level)),
                       DV = rnorm(num_levels * num_per_level, 0, 1))
    anova_out <- aov(DV ~ level, sim_data)
    f_ <- summary(anova_out)[[1]]$`F value`[1]
    f_values[i] <- f_
    
    f_values_df <- tibble(f_value = f_values,
                          sample = 1:length(f_values))
  }
  return(f_values_df)
}

sim_f_values <- gen_sim_f(levels, n_per_level, n_samples)

ggplot(sim_f_values) +
  geom_histogram(aes(x = f_value), bins = 100) +
  geom_vline(aes(xintercept = 4.737, color = "red"), linetype = "dashed") +
  annotate(x = 4.737, y = 50, label = "Critical Value = 4.737", geom = "label")
  
```

### Creating a function that adds an effect to the first level. This will allow for the design to produce an F-value that is smaller or larger than the critical F-value.

```{r}
gen_effect_data <- function(num_levels, num_per_level, num_samples, effect_size) {
  
  column_names = c("subjects", "IV", "DV", "sample")
  
  total_sim_data <- as_tibble(t(column_names))[0, ]
  
  for(i in 1:num_samples) {
    sim_data <- tibble(subject = 1:(num_levels * num_per_level),
                       level = as.factor(rep(1:num_levels, each = num_per_level)),
                       DV = rnorm(num_levels * num_per_level, 0, 1),
                       sample = i)
    
    sim_data[sim_data$level == 1, ]$DV <- sim_data[sim_data$level == 1, ]$DV + effect_size
    
    total_sim_data <- rbind(total_sim_data, sim_data)
    
  }
  
  total_sim_data <- total_sim_data %>%
    group_by(subject, level) %>%
    summarise(DV = mean(DV))
  
  return(total_sim_data)
}
```

## (1) 

Create simulated data for the above design that could be produced by the null hypothesis, and that results in a 𝐹 value that is smaller than the critical value for 𝐹 in this design (assume alpha = .05). Report the ANOVA, and show a ggplot of the means in the simulated data. Furthermore, display the individual data points on top of the means. Would you reject the null hypothesis in this situation, and would you be incorrect or correct in rejecting the null?

### Creating simulated data with an effect size applied to level 1 of 0.01 and displaying the ANOVA summary.

```{r}
lower_effect_data <- gen_effect_data(levels, n_per_level, n_samples, 0.01)

lower_effect_anova <- aov(DV ~ level, lower_effect_data)

unclass(summary(lower_effect_anova)) %>%
  kbl(digits = 2) %>%
  kable_classic(position = "center")
```

### Graphing the F value against the null distribution.

```{r}
lower_f_value <- summary(lower_effect_anova)[[1]]$`F value`[1]

ggplot(sim_f_values) +
  geom_histogram(aes(x = f_value), bins = 100) +
  geom_vline(aes(xintercept = 4.737, color = "red"), linetype = "dashed") +
  annotate(x = 4.737, y = 50, label = "Critical Value = 4.737", geom = "label") +
  geom_vline(aes(xintercept = lower_f_value, color = "blue")) +
  annotate(x = lower_f_value, y = 75, xend = 0, label =  paste0("Simulated F value = ", round(lower_f_value, 2)), geom = "label")
```

### Graphing the means of the simulated data where the effect size was 0.01.

```{r}
lower_effect_data_grouped <- lower_effect_data %>%
  group_by(level) %>%
  summarise(mean_DV = mean(DV),
            SE = sd(DV)/length(DV))

ggplot(lower_effect_data_grouped, aes(x = level, y = mean_DV)) +
  geom_bar(stat = "identity", aes(fill = level)) +
  scale_y_continuous(limits = c(-0.05, 0.05)) +
  geom_errorbar(aes(ymin=mean_DV - SE,
                    ymax=mean_DV + SE), width=.1) + 
  geom_point(data = lower_effect_data, mapping = aes(x = level, y = DV), alpha = 0.5) +
  geom_point(alpha=.25) +
  ylab("DV")
```

I would not reject the null in this situation as the F value calculated from the simulated data is less than the critical value. The parameters of the simulation, however, can produce data that leads to an F value greater than the critical value, but that is unlikely.

## (2) 

Create simulated data for the above design that could be produced by the null hypothesis, and that results in a 𝐹 value that is larger than the critical value for 𝐹 in this design (assume alpha = .05). Report the ANOVA, and show a ggplot of the means in the simulated data. Furthermore, display the individual data points on top of the means. Would you reject the null hypothesis in this situation, and would you be incorrect or correct in rejecting the null?

### Creating simulated data with an effect size applied to level 1 of 0.035 and displaying the ANOVA summary.

```{r}
higher_effect_data <- gen_effect_data(levels, n_per_level, n_samples, 0.035)

higher_effect_anova <- aov(DV ~ level, higher_effect_data)

unclass(summary(higher_effect_anova)) %>%
  kbl(digits = 2) %>%
  kable_classic(position = "center")
```

### Graphing the F value against the null distribution.

```{r}
higher_f_value <- summary(higher_effect_anova)[[1]]$`F value`[1]

ggplot(sim_f_values) +
  geom_histogram(aes(x = f_value), bins = 100) +
  geom_vline(aes(xintercept = 4.737, color = "red"), linetype = "dashed") +
  annotate(x = 4.737, y = 50, label = "Critical Value = 4.737", geom = "label") +
  geom_vline(aes(xintercept = higher_f_value, color = "blue")) +
  annotate(x = higher_f_value, y = 75, xend = 8.0, label =  paste0("Simulated F value = ", round(higher_f_value, 2)), geom = "label") +
  expand_limits(x = 10)
```

### Graphing the means of the simulated data where the effect size was 0.01.

```{r}
higher_effect_data_grouped <- higher_effect_data %>%
  group_by(level) %>%
  summarise(mean_DV = mean(DV),
            SE = sd(DV)/length(DV))

ggplot(higher_effect_data_grouped, aes(x = level, y = mean_DV)) +
  geom_bar(stat = "identity", aes(fill = level)) +
  scale_y_continuous(limits = c(-0.05, 0.05)) +
  geom_errorbar(aes(ymin=mean_DV - SE,
                    ymax=mean_DV + SE), width=.1) + 
  geom_point(data = higher_effect_data, mapping = aes(x = level, y = DV), alpha = 0.5) +
  geom_point(alpha=.25)
```

I would reject the null in this situation as the F value calculated from the simulated data is greater than the critical value. The parameters of the simulation, however, can produce data that leads to an F value less than the critical value, but that is unlikely.

## Bonus Question: 

In the lab we saw that F-distribution is robust to violations of the assumptions of ANOVA. For example, the simulation of the null based on a bi-modal distribution was very similar to the true F distribution. For this bonus question, show that you can “break” the F-distribution. Specifically, can you run a simulation that samples numbers from a non-normal distribution that does produce a very different looking F-distribution?

### Breaking the F-distribution by sampling data from a Weibull distribution (shape = 0.1, scale = 1000).

```{r}
gen_sim_f <- function(num_levels, num_per_level, num_samples) {
  f_values <- length(num_samples)

  for(i in 1:num_samples) {
    sim_data <- tibble(subject = 1:(num_levels * num_per_level),
                       level = as.factor(rep(1:num_levels, each = num_per_level)),
                       DV = rweibull(num_levels * num_per_level, shape = 0.1, scale = 1000))
    anova_out <- aov(DV ~ level, sim_data)
    f_ <- summary(anova_out)[[1]]$`F value`[1]
    f_values[i] <- f_
    
    f_values_df <- tibble(f_value = f_values,
                          sample = 1:length(f_values))
  }
  return(f_values_df)
}

sim_f_values <- gen_sim_f(levels, n_per_level, n_samples)

ggplot(sim_f_values) +
  geom_histogram(aes(x = f_value), bins = 100)
  
```