---
title: "Lab 6 - Comparisons"
author: "Drew Shives"
date: "2022-01-30"
slug: lab-6-comparisons
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2022-04-27T13:31:58-05:00'
featured: no
type: book
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (1) Section 12.3.3 from your textbook refers to: The problem with replications of a meaningless experiment: ‘alpha and the captain’s age’. The issue here is that if you run an ineffectual experiment enough times you can always find a significant result by chance. The textbook mentions that if you repeat an experiment 20 times, you are guaranteed to find a significant result with .64 probability, and the probability is .92 if you repeat the experiment 50 times.

### (1a) Make use of the rbinom() function to show you can reproduce both probabilities. 

```{r}
a_20 <- rbinom(10000, 20, 0.05)

hist(a_20)

length(a_20[a_20 > 0]) / 10000

```

```{r}
a_50 <- rbinom(10000, 50, 0.05)

hist(a_50)

length(a_50[a_50 > 0]) / 10000

```

### (1b) If the ineffectual experiment was conducted 20 times, and there were four groups, and the experimenter would accept a significant result from any of the orthogonal linear contrasts, what would be the probability of finding a significant result here? 

```{r}

B <- replicate(10000, sum(rbinom(20, 3, 0.05)))

length(B[B > 0]) / 10000

```

## (2) Consider that a researcher publishes a study showing a significant effect, p <. 05; but, in reality the researcher makes a type I error, and the manipulation did not cause any difference. If many other researchers replicated the study, what kind of p-values would they find? Use R to create a sampling distribution of p-values that would be expected in this situation. What shape does this distribution have?

```{r}
p_values <- replicate(10000, t.test(rnorm(10, 0, 1), rnorm(10, 0, 1), var.equal = TRUE)$p.value)

hist(p_values)
```

## (3) Now assume that the published result reflects a true effect. Specifically, let’s imagine the study had two groups (between-subjects), with 20 subjects in each group. Assume that scores for subjects are all sampled from a normal distribution, and that group A has larger mean than group B by .5 standard deviations (e.g., Cohen’s d = .5). If many other researchers replicated the study, what kind of p-values would they find? Use R to create a sampling distribution of p-values that would be expected in this situation. What shape does this distribution have?

```{r}
p_values <- replicate(10000, t.test(rnorm(20, 0, 1), rnorm(20, 0.5, 1), var.equal = TRUE)$p.value)

hist(p_values)
```


