---
title: "Lab 7 - Factorial ANOVA"
author: "Drew Shives"
date: "2022-01-29"
slug: lab-7-factorial-anova
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2022-04-27T13:31:58-05:00'
featured: no
type: book
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
library(tidyverse)
```

## (3) In chapter 10 of Crump et al. (2018), there is a discussion of patterns of main effects and interactions that can occur in a 2x2 design, which represents perhaps the simplest factorial design. There are 8 possible outcomes discussed https://crumplab.github.io/statistics/more-on-factorial-designs.html#looking-at-main-effects-and-interactions. Examples of these 8 outcomes are shown in two figures, one with bar graphs, and one with line graphs. Reproduce either of these figures using ggplot2.

```{r}
pattern1 <- tibble(IV1 = c("A","A","B","B"),
                IV2 = c("1","2","1","2"),
                means = c(6,6,6,6))

pattern2 <- tibble(IV1 = c("A","A","B","B"),
                IV2 = c("1","2","1","2"),
                means = c(10,10,5,5))

pattern3 <- tibble(IV1 = c("A","A","B","B"),
                IV2 = c("1","2","1","2"),
                means = c(10,13,5,2))

pattern4 <- tibble(IV1 = c("A","A","B","B"),
                IV2 = c("1","2","1","2"),
                means = c(5,10,10,15))

pattern5 <- tibble(IV1 = c("A","A","B","B"),
                IV2 = c("1","2","1","2"),
                means = c(10,18,5,7))

pattern6 <- tibble(IV1 = c("A","A","B","B"),
                IV2 = c("1","2","1","2"),
                means = c(10,2,10,2))

pattern7 <- tibble(IV1 = c("A","A","B","B"),
                IV2 = c("1","2","1","2"),
                means = c(2,12,5,9))

pattern8 <- tibble(IV1 = c("A","A","B","B"),
                IV2 = c("1","2","1","2"),
                means = c(5,10,10,5))

all <- rbind(pattern1, pattern2, pattern3, pattern4, pattern5, pattern6, pattern7, pattern8)
type <- c(rep("~1, ~2, ~1x2",4),
          rep("1, ~2, ~1x2",4),
          rep("1, ~2, 1x2",4),
          rep("1, 2, ~1x2",4),
          rep("1, 2, 1x2",4),
          rep("~1, 2, ~1x2",4),
          rep("~1, 2, 1x2",4),
          rep("~1, ~2, 1x2",4))
type<-as.factor(type)

all <- cbind(all,type)

ggplot(all, aes(x=IV1, y=means, group=IV2, fill=IV2))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  facet_wrap(~type, nrow=2)+
  theme(legend.position = "top")

ggplot(all, aes(x=IV1, y=means, group=IV2, color=IV2))+
  geom_point()+
  geom_line()+
  theme_classic()+
  facet_wrap(~type, nrow=2)+
  theme(legend.position = "top")
```

## (4) In the conceptual section of this lab we used an R simulation to find the family-wise type I error rate for a simple factorial design with 2 independent variables. Use an R simulation to find the family-wise type I error rate for a factorial design with 3 independent variables.

```{r}
simulation <- rbinom(10000,7,.05)
length(simulation[simulation > 0])/10000

simulation_save <- tibble()

for(i in 1:10000){
  n <- 12
  factorial_data <- tibble(A = factor(rep(c("L1","L2"), each = n)),
                         B = factor(rep(rep(c("L1","L2"), each = n/2),2)),
                         C = factor(rep(c("L1","L2"), n)),
                         DV = rnorm(n*2,0,1))

  output <- summary(aov(DV ~ A*B*C, data = factorial_data))

  simulation_tibble <- tibble(p_values = output[[1]]$`Pr(>F)`[1:7],
                       effect = c("A","B","C","AxB","AxC","BxC","AxBxC"),
                       simulation = rep(i,7))

  simulation_save <-rbind(simulation_save, simulation_tibble)
}

type_I_errors <- simulation_save %>%
  filter(p_values < .05) %>%
  group_by(simulation) %>%
  count()

dim(type_I_errors)[1]/10000
```






