---
title: "Lab 8 - Contrast Analyses"
author: "Drew Shives"
date: "2022-01-28"
slug: lab-8-contrast-analyses
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2022-04-27T13:31:58-05:00'
featured: no
type: book
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(usefun)
```

## (1) Consider a 2x2 design. Assume the DV is measured from a normal distribution with mean 0, and standard deviation 1. Assume that the main effect of A causes a total shift of .5 standard deviations of the mean between the levels. Assume that level 1 of B is a control, where you expect to measure the standard effect of A. Assume that level 2 of B is an experimental factor intended to reduce the effect of A by .25 standard deviations.

### (1a) Create a ggplot2 figure that depicts the expected results from this design.

```{r}

grand_mean <- 0
A <- c(0,.5)
B <- c(0,0)
AB <- c(0,0,0,-.25)

model_data <- tibble()

for(i in 1:length(A)){
  for(j in 1:length(B)){
    IVA <- i 
    IVB <- j
    DV <- grand_mean + A[i] + B[j] + AB[(i-1) * length(B) + j]
    sc_GM <- grand_mean
    sc_A <- A[i]
    sc_B <- B[j]
    sc_AB <- AB[(i-1) * length(B) + j] 
    row_entry <- tibble(IVA,IVB,DV,
                        sc_GM,sc_A,sc_B,sc_AB)
    model_data <- rbind(model_data,row_entry)
  }
}

knitr::kable(model_data)

bar_graph <- ggplot(model_data, 
                    aes(y=DV,
                        x=as.factor(IVA),
                        fill=as.factor(IVB))) + 
  geom_bar(stat='identity', position='dodge')

line_graph <- ggplot(model_data, 
                     aes(y=DV,
                         x=IVA,
                         linetype=as.factor(IVB))) + 
  geom_line() + 
  geom_point()

bar_graph

line_graph

```

### (1b) How many subjects are needed to detect the main effect of A with power = .8?

```{r}
power_function <- function(num_subjects, effect) {
  
  p_value <- c()
  
  effect_num = c()
  
  if(effect == "A") {
    effect_num = 1
  } else if(effect == "B") {
    effect_num = 2
  } else if(effect == "AB") {
    effect_num =3
  }

  for(i in 1:1000){
    
    simulation_df <- tibble(
      IVA = rep(rep(c("1","2"), each = 2), num_subjects),
      IVB = rep(rep(c("1","2"), 2), num_subjects),
      DV = rnorm(4 * num_subjects, c(0,0,.5,.25), 1)
    )
    aov_results <- summary(aov(DV ~ IVA * IVB, simulation_df))
    p_value[i] <- aov_results[[1]]$`Pr(>F)`[effect_num]
  }
  
  length(p_value[p_value < 0.05]) / 1000
}

```

```{r}
n <- c(seq(50, 60))


for(i in n) {
  print(i)
  print(power_function(i, "A"))
  print_empty_line()
}

```

### (1c) How many subjects are needed to detect the interaction effect with power = .8?

```{r}
n <- c(400, 425, 450, 475, 500, 525, 550, 575, 600)


for(i in n) {
  print(i)
  print(power_function(i, "AB"))
  print_empty_line()
}

```


