---
title: "R Basics"
author: "Drew Shives"
date: "2022-01-31"
slug: r-basics
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2022-04-27T13:31:58-05:00'
featured: no
type: book
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
library(tidyverse)
```

## (1) Different object types, including vectors and data.frames. Also, demonstrate your knowledge of at least two or more object types that are not vectors or data.frames.

```{r}
# Vectors

vector1 <- c(1, 2, 3, 4, 5)
vector2 <- c("a", "b", "c", "d", "e")

# Lists

list1 <- list(1, 2, 3, 4, 5)
list2 <- list("a", "b", "c", "d", "e")

# Tibbles

tibble1 <- tibble(
  numbers = c(1, 2, 3, 4, 5),
  letters = c("a", "b", "c", "d", "e")
)

# Matrices

matrix1 <- matrix(data = tibble1)

# Dataframes

df1 <- data.frame(
  numbers = c(1, 2, 3, 4, 5),
  letters = c("a", "b", "c", "d", "e")
)
```

## (2) Loops. There are multiple ways to iterate in R. Demonstrate at least two ways, including a FOR loop. Show that you know how to break a loop. Highlight one thing about loops in R that was not obvious to you, or you think might not be obvious to someone else.

```{r}
# For loops

for(i in 1:5) {
  print(i)
}

# While loop

loop_var = 0

while(loop_var <= 5) {
  print(loop_var)
  loop_var = loop_var + 1
}

# I think one thing about loops that is not initially obvious is their utility. The ability to iterate over a variable (or two+ variables if using the foreach library) is incredibly useful when trying to create data, run a function multiple times, etc.

```

## (3) Logic. Show that you understand logical operations in R. Show that you know how to use logical operations for indexing. For example, use logical operations to index a vector, and to index a data.frame.

```{r}
# If statements

if_var <- sample(c("a", "b", "c", "d"), 1)

if(if_var == "a") {
  print(vector1[1])
} else if(if_var == "b") {
  print(vector1[2])
} else if(if_var == "c") {
  print(tibble1[3, ])
} else {
  print(tibble1[4, ])
}

```

## (4) Loops and logic. Demonstrate that you can combine loops and logic to solve some problem. For example, use loops and logic to solve the fizz buzz problem.

```{r}
# Loops and logic: For loop with an if statement

for(i in sample(c(1:5), 1)) {
  if(i <= 3) {
    print("low")
    break
  } else {
    print("high")
    break
  }
}
```

## (5) Pick at least one advanced problem from this list of problems, and show that you can solve it in R. - SNAKES AND LADDERS

```{r}
# Ladders:
# 3 goes to 11
# 6 goes to 17
# 10 goes to 12
# 9 goes to 16

# Snakes:
# 14 goes to 4
# 19 goes to 8
# 22 goes to 20
# 24 goes to 16


total_rolls <- c()

for(i in 1:1000) {
  
  end_state <- 0
  num_rolls <- 0
  
  while(end_state < 25) {
    
    num_rolls <- num_rolls + 1
    roll <- sample(1:6, 1)
    end_state <- end_state + roll
    
    if(end_state == 3) {
      end_state = 11
    } else if (end_state == 6) {
      end_state = 17
    } else if (end_state == 10) {
      end_state = 12
    } else if (end_state == 9) {
      end_state = 16
    } else if (end_state == 14) {
      end_state = 4
    } else if (end_state == 19) {
      end_state = 8
    } else if (end_state == 22) {
      end_state = 20
    } else if (end_state == 24) {
      end_state = 16
    }
  }
  total_rolls <- append(total_rolls, num_rolls)
}

mean(total_rolls)
```

```{r}
means <- c()

for(i in 1:1000) {
  total_rolls <- c()
  
  for(i in 1:1000) {
    
    end_state <- 0
    num_rolls <- 0
    
    while(end_state < 25) {
      
      num_rolls <- num_rolls + 1
      roll <- sample(1:6, 1)
      end_state <- end_state + roll
      
      if(end_state == 3) {
        end_state = 11
      } else if (end_state == 6) {
        end_state = 17
      } else if (end_state == 10) {
        end_state = 12
      } else if (end_state == 9) {
        end_state = 16
      } else if (end_state == 14) {
        end_state = 4
      } else if (end_state == 19) {
        end_state = 8
      } else if (end_state == 22) {
        end_state = 20
      } else if (end_state == 24) {
        end_state = 16
      }
    }
    total_rolls <- append(total_rolls, num_rolls)
  }
  
  means <- append(means, mean(total_rolls))
}

hist(means)

```













